How has the Brazilian Amazon been constructed as a problem? Presidential speeches and transnational politics since 1985

### 2023-02-01 Submission/Re-submission/draft/draft_resubmission.rmd
all changes into markdown

```{=latex}
\keywords{ discourse analysis, transnational governance, environmental policy, Brazilian Amazon, supervised learning, deforestation}
```

```{=latex}
\wordcount{ 7685 (including abstract, text, references, and footnotes)}
```

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(include = FALSE)
library(dplyr)
library(tidyr)
library(scales)
library(RColorBrewer)
library(stargazer)
library(kableExtra)
library(tibble)
library(readxl)
library(ggplot2)
library(gridExtra)
library(modelsummary)
library(plm)
Sys.setlocale("LC_ALL", "C")
options(stringsAsFactors = FALSE, scipen = 999)
```

@@

*We need to protect the Amazon from foreign interests. We need to exploit the Amazon's natural resources. We need to provide better living standards for the people in the Amazon. We need to preserve the Amazon as a standing ecosystem*. Each of these statements contains an implicit assumption of what needs to be solved, or in other words, it constructs the Amazon as a particular problem. Scholars studying the Brazilian Amazon often identify different governments throughout Brazilian history as proponents of a view of the region, its people, and forest [@drummond2006; @padua2012; @franchini2019; @capobianco2019; @pereira2021]. While the military dictatorship (1964-85) is, for example, associated with understanding the Amazon as issues of national sovereignty and economic integration, Sarney's (1985-1990) and Lula's (2003-2010) presidencies are often tied to environmental conservation. Yet, such monolithic representations advance the view that governments understood and acted consistently towards the Amazon as a specific problem, masking government diversity of discursive constructions and policies.

In this article, we show how the Brazilian Amazon has been constructed as a problem transnationally in presidential discourses since 1985. Building on @bacchi2009's theory of problem-representation and @keck1998's theory of transnationalism, we offer a framework that ties discursive problem issue representations to relationships that transcend nation-states to identify and explain variation in problems-constructions. We shift the focus from governmental policy to discourse to capture a broader range of understandings for policy objects, problems that demand dedicated policy attention, that introduce, shape, and justify outcomes to broad audiences [@zarefsky2004; @gillion2016]. Although problem-constructions take place in a series of instances, we analyze the case of official presidential discourses. The Brazilian executive branch of government, led by the president, has historically been the primary actor in shaping public policy in Brazil [@pinto2016]. Working with presidential discourses also allows us to identify variations in how the Amazon is discursively constructed in transnational politics beyond environmental policies. Whereas understudied in the Amazonian context, presidential speeches legitimize perceptions about the Amazon that are key for policy adoption and implementation [see @alesina2009; @lopez2020]. This is especially pertinent for deforestation in the Amazon as expectations of local actors, generated from material and discursive governmental practices, are crucial in decisions to deforest on the ground [@assuncao2015; @capobianco2019; @capobianco2021; @campbell2015]. As crucial participants in transnational politics, presidents deploy different problem-constructions according to the transnational audiences they are speaking to.

To investigate how the Brazilian Amazon has been constructed as a problem in presidential speeches, we create a dataset containing 6240 official presidential speeches by all Brazilian presidents since 1985. Next, we identify all instances where presidents refer to the Amazon in their speeches. We develop a codebook grounded in Amazonian historiography to code how each of these statements constructs Amazon as a particular problem. With this codebook, we manually code a randomly selected training set of statements and use supervised machine-learning to label the remaining observations. We then conduct a descriptive analysis of when, where, and how presidents construct the Amazon as a problem.

Our findings are threefold. First, we find that the frequency in which the Amazon appears as a topic in presidential speeches is driven by domestic and international environmentally-related events. Second, we find that presidents are more likely to construct the Amazon as an issue of environmental conservation outside of the Amazon region, this is especially true when presidents speak internationally. Within the Amazon region, presidents tend to construct the Amazon as an issue of social development and economic integration. Third, the same government adopts a diverse range of problem-construction. While the constructions of the Amazon as an issue of economic integration were more salient from 1985 to the mid-2000s, environmental conservation and social development problem-constructions steadily grew since the 2000s, temporarily surpassing economic integration as a more salient construction from 2010 to 2015. Constructing the Amazon as an issue of sovereignty steadily increased after 2010. This article provides the first comprehensive overview of when, where, and how the Amazon is constructed as problems in presidential speeches.

# 2. Theory: problem-constructions, transnational settings, and presidential speeches

## 2.1 Governments and problem-constructions

@@

@@

We argue that the reason behind these monolithic representations of governments is conceptual: many studies of the Brazilian Amazon contain an explicit, or implicit assumption, of what the Amazon problem is. @bacchi2009 [p. 10] posits that policies have a cultural dimension as they take "shape within specific historical and national or international contexts". The existence or proposal of a policy implies that there is a (public) problem that needs (governmental) action to be fixed. The alleged problem is not always explicitly stated in a policy, but policies are represented as solutions for both explicit and implicit problems. If the Amazon is represented as an "environmental" problem, as is often the case in the literature, then it demands an "environmental" solution. Consequently, studies focus on understanding environmental policies [@capobianco2019], the work of environmental NGOs [@silva-muller2022], or the rise of global environmental governance [@andonova2022]. Alternatively, studies looking into the developments of social and fiscal policy in the region during the last decades are scarce [@acker2017]. We acknowledge the quality and importance of all these studies. Still, we argue that it is crucial for scholars to study the Amazon as a biological, social, political, and geographical policy object. We conceptualize the Amazon as a policy object that demands dedicated policy attention of a specific nature. Hence, governments can highlight diverse explicit and implicit problems related to a policy object that need political attention. This is what we label as problem-constructions. <!-- Are we missing a transition from problem-constructions generally to the case of speeches? -->

@@

The possibility of different problem-constructions for the same policy object in political discourses, requires an explanation of why they would vary. On the one hand, @putnam1988 theory of two-level games in international negotiations can be adapted to explain variation in presidential speeches. @putnam1988 argues that the outcomes of international negotiations lie within the overlap between the interests (and pressure) of domestic and international groups. At the national level, interest groups pressure the government to adopt favorable policies. Internationally, the government seeks to fulfill domestic preferences "while minimizing the adverse consequences of foreign developments" [@putnam1988, p. 434]. While the theory has been expanded to cover multi-levels games [see @collinson1999], its core remains the same: the outcome of negotiations should lie within the overlap of diverse interest groups.

@@

On the other hand, @keck1998 theory of international politics as transnational networks, developed in the context of value-based advocacy groups, provides an alternative explanation for variation in problem-constructions. The authors reject the dichotomy of international and domestic levels arguing that the identity and goals of state and non-state actors are not derived from their structural location vis-a-vis domestic and foreign interest groups. Rather, international politics as transnational networks have "a structured and structuring dimension" [@keck1998, p. 4]: state and non-state actors participate in and shape international politics. In this sense, the positions state actors as components of a transnational network derive from complex interactions in which their identities and goals that are malleable and can vary across time and space.

Important participants of transnational networks, as presidents, have a myriad of tactics to introduce, define, persuade, socialize, and pressure when interacting with other actors. Presidents can, for example, engage in information politics by creating politically meaningful information and/or moving it where they believe it can have an impact [@keck1998, p. 24]. As well, presidents can engage in leverage politics by leveraging institutions to realize symbolic or material gains. Transnational networks do not assume a consistent common denominator of problem-constructions. That is, presidents construct problems and attempt to inform or leverage transnational audiences for symbolic or material gains according to their expectations of said audience. This indicates that how presidents construct policy objects as problems can vary within the country (i.e. speaking at diverse states or different associations) as well as when speaking internationally (i.e. during a climate conference or after a catastrophe). We call these levels in which there could be variation in how presidents construct a problem a setting. The setting is a level of aggregation that encompasses one, or multiple, audiences and makes sense concerning the policy object under study, in this case, the Amazon.

## 2.3 Problem-constructions and policy

In addition to the setting where the speech takes place, it is natural that outcomes in policy adoption and implementation can affect when and how presidents construct problems. @gillion2016 argues that there are two types of outcomes that can affect presidential discourses: institutional and societal. An institutional outcome relates to how presidential discourses can shape policy coalitions and the policy agenda. While the relationship between presidential speeches and policy may not be causal, they have the power to introduce, define, and justify public policy, as well as shape issue perception to broad audiences [@zarefsky2004]. This is particularly true in the Brazilian context where the executive branch of government, led by the president, have historically been the primary players in shaping public policy, especially after the 1988 constitution [@pinto2016].

A societal outcome, instead, relates to how discourse informs attitudes and behavior of people on the ground and affects policy outcomes. In the case of deforestation, for example, Amazonian scholars argue that the expectations of local actors generated from governments' decisions [@assuncao2015] and discourses [@capobianco2019; @capobianco2021; @campbell2015] are a crucial factor in choice to deforest on the ground. As presidents participate and shape transnational politics by constructing the Amazon as a certain issue, discourses define policy agendas, direct "political officials' at multiple stages of the public policy-making process" [@gillion2016, 19], influence people's decision on the ground, and justify previous decision or policy outcomes.

Considering the possibility of varied problem-construction, the relationship between policy and discourse becomes endogenous and co-constituted: problem constructions in presidential speeches can justify past outcomes and/or lead to future outcomes. A bad track record in climate change policy outcomes can be justified with a good track record in energy security. This was, for example, the strategy adopted by President Bush during the Kyoto Protocol negotiations [@calderwood2020]. As presidents participate in transnational networks, with malleable identities and goals, will refer to outcomes interchangeably, based on their expectations about the setting they are speaking at and the transnational audiences present. Policy outcomes, therefore, can be (re)constructed in presidential discourse in diverse forms to shape certain problem definitions and possible solutions.

@@

## 3.1 Amazonian historiography

To capture the cultural dimension and historical contexts behind problem constructions, we rely on the Amazonian historiography literature. We refer to Amazonian historiography as the body of research conducted by social and environmental scientists that tell the story of diverse governmental programs that studies the Amazon region, forest, and its people. Within this literature, we identify three historical-political constructions: national sovereignty, economic integration, and environmental conservation. Based on an inductive analysis of our empirical material, we argue that a fourth separate construction also exists: social development. The social development construction focuses on the constitutional rights and dignity of the inhabitants of the Amazonian region.

### National Sovereignty

@@

### Economic Integration

@@

### Environmental Conservation

The rapid economic changes in the region from the 1960s to the 1980s were matched with the birth of environmental institutions [@drummond2006]. Arguably, the creation of these institutions stemmed from the impression of the lack of control over the environment following the efforts to economically integrate in the region [@acker2021; @capobianco2021; @hecht1990]. This process accelerated in the late 1980s with the birth of modern environmentalism [@viola1987], epitomized in the 1992 Earth Summit in Rio de Janeiro [@hochstetler2021; @capobianco2021]. The environmental conservation problem-construction emphasizes that the Amazon should be preserved, deforestation should be halted, and the sustainable practices of indigenous and local peoples should be maintained through the protection of their territories and rights to self-determination [@hochstetler2007]. The policy solution implies, for example, further investments in command-and-control institutions, the valuation of standing ecosystems through incentive schemes, and the creation of protected areas.

### Social Development

@@

## 3.2 Data and Operationalization

To identify varied problem-constructions, we focus on presidential speeches. By focusing on presidential speeches, rather than specific policy domains, we analytically avoid hinging our analysis on an underlying assumption of what the problem is. Thus, we are not necessarily prioritizing an environmental, economic, or sovereignty-based view of the Amazon. We build upon @cezar2020 dataset, which contains all official speeches by Brazilian Presidents from 1985 to 2019. We update the dataset by scraping and adding all official speeches from 2020 to 2022. The final dataset encompasses 6240 speeches. We proceed to identify speeches that refer to the Amazon as a region, people, or forest. We do so by detecting all speeches in which the stem "amazon" appears. In Portuguese, the stem captures terms such as "Amazonia", "Amazonica", "Amazonidas", "Amazonense(s)", "Amazonas", among others. We find that 968 speeches are, at least partially, about the Amazon. Using the poldis R package [@sposito2021], we extract two sentences before and after the sentence in which the stem "amazon" appears. We opt for picking two sentences around, rather than words, because sentences usually contain a cohesive idea. This is how our unit of analysis, Amazonian statements, are generated. This process yields 2048 unique Amazonian statements about the Amazon. On average, an Amazonian statement contains 131 words.

We code Amazonian statements in three steps. First, we develop a codebook for coding Amazonian problem-constructions based on the Amazonian historiography (see Codebook in appendix). In their conceptualization, each problem-construction is mutually exclusive; that is, they cover different forms of constructing the Amazon as a problem. However, an Amazonian statement can be coded as referring to multiple problem constructions (see Figure 1 below). Amazonian statements, thus, can be pure-type statements (i.e. constructs the Amazon as a single problem) or mixed-types (i.e. constructs the Amazon as multiple issues).

@@

Second, with the codebook in hand, each one of the authors separately hand-coded 1024 randomly selected Amazonian statements. This amount refers to 50% of all the Amazonian statements identified. We chose to hand code half of the observations because there are several nuances in how presidents talk about the Amazon over time. As well, since we have four different code categories, the size of the training set should be increased [see @grimmer2022]. The hand-coding strategy provides a robust and reliable number of observations to be used for training and validation of the automated model. Once hand coding was completed, the average intercoder agreement for hand-coded categories was 85%. For each non-matching coded observation, the authors discussed and sorted their disagreements, which were subsequently specified in the codebook.

Third, the hand-coded data is randomly divided into a training set, containing 80% of the data (819 observations), and a validation set, containing the remaining 20% of the data (205 observations). We employ a support-vector machine (SVM) algorithm, a non-probabilistic linear classifier appropriate for binary categories, to label texts [@meyer2021; @noble2006]. To validate the model, the SVM algorithm is trained using the hand-coded training set and employed to classify observations in the validation set. As a final robustness check, the authors randomly select 100 automatically coded statements and find that the automated model is 95% accurate in coding these statements as the authors would. Automating the coding of the remaining half of the observations is estimated to have saved the authors over one month of work in comparison to manual coding. The final dataset for analysis, excluding false positive matches, contains 1924 coded Amazonian statements.

## 3.3 Analysis and Limitations

To analyze our data, we first visualize when presidents speak more about the Amazon and how they construct the Amazon as a problem in time. We then model the share of a specific pure problem-construction, as the dependent variable, relative to the settings in which a particular speech happens, as our independent variable. This means we predict the probability of a president constructing the Amazon as an issue of (1) environmental conservation, (2) economic integration, (3) social development, (4) national sovereignty, separately, and in relation to the remaining problem-constructions (see table 1 below). We focus on single problem constructions to provide unambiguous interpretations and more robust findings. We divide settings in terms of non-Amazonian states, Amazonian states, Amazonian countries, and non-Amazonian countries. This level of aggregation refers to geo-political settings that make sense in light of the Amazon, as our policy object. This aggregation combines different expectations presidents can have about transnational audiences in each of these settings. The model controls for yearly deforestation rates, inflation, election years, and peak years for Amazon in presidential speeches.

```{r table 1, fig.cap="Models Description", echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE, include=TRUE}

## Count average words in Amazonian Statements
## Need to use original text data
# amazon_speeches_long <- readRDS("~/Documents/GitHub/amazondef/Replication_Files/amazon_speeches_long.Rds")
# sum(stringr::str_count(amazon_speeches_long$AM2, '\\w+'))/2048
## 131 on average

models <- tibble::tribble(~"", ~Description,
                          "Model", "",
                          "Fixed-effects logistic regressions indexed by the president", "The regression method allows to control for variables not measured (e.g. ideology) related to individuals (i.e. presidents) [see @allison2009].",
                          "Dependent variables", "",
                          "Problem constructions", "Pure environmental conservation; pure economic integration; pure social development; and pure national sovereignty.",
                          "Independent variable", "",
                          "Settings", "Setting where speech occurs as (1) non-Amazonian Brazilian states, (2) Brasilia, (3) Amazonian countries, and (4) international, in relation to the reference category of Amazonian states within Brazil.",
                          "Control variables", "",
                          "Deforestation rates", "Lagged yearly deforestation rates as a control for the possible relationship between deforestation rates and Amazonian problem-constructions.",
                          "Election year", "Election years are controls for possible discursive changes related to the Amazon for these years.",
                          "Inflation rate",  "Average yearly inflation rates as a control for a possible relationship between economic conditions/fluctuations and problem constructions.",
                          "High-profile environmental event", "Years in which high-profile environmental events occurred control for the frequency and, possibly, how the Amazon was constructed as a problem in discourse for those years.")

kbl(models, booktabs = T, caption = "Model description") %>%
kable_styling(latex_options = c("striped", "repeat_header", "scale_down"),position= "center", font_size = 9, full_width = TRUE) %>% row_spec(c(1,3,5,7), bold = TRUE, italic = TRUE)
# see also: pack_rows("Model", 1, 2)
# https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_pdf.pdf
```

The main limitations of our approach are threefold. First, related to our subsetting and coding strategies, we classify statements as Amazonian based on a dictionary composed of a single lexicon stem: "amazon". We chose to do so knowing that a few speeches about the Amazon might not contain the lexicon "amazon", for example, when the president says, "the forest" or "deforestation". However, we consider this safer as we cannot be sure that mentions of the forest or deforestation do not correspond to other biomes such as the Cerrado or the Atlantic Forest.

>>>>

```{=latex}
\keywords{ discourse analysis, transnational governance, environmental policy, Brazilian Amazon, supervised learning, deforestation}
```
```{=latex}
\wordcount{ 7685 (including abstract, text, references, and footnotes)}
```
```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(include = FALSE)
library(dplyr)
library(tidyr)
library(scales)
library(RColorBrewer)
library(stargazer)
library(kableExtra)
library(tibble)
library(readxl)
library(ggplot2)
library(gridExtra)
library(modelsummary)
library(plm)
Sys.setlocale("LC_ALL", "C")
options(stringsAsFactors = FALSE, scipen = 999)
```

@@

> *"[...] I will fight against the NGOs on this issue of associating the increase in deforestation to our agricultural expansion. First of all, these NGOs should go plant trees in their own countries [...]" (President Lula, Cuiaba, 2007).*

> *[...] It is essential that, when taking care of the environment, we do not overlook people: the Amazon region is home to more than 20 million inhabitants, including indigenous peoples and riverside dwellers [...]" (President Bolsonaro, New York City, 2022)*

Each of the above statements contains an implicit or explicit assumption of what needs to be solved, or in other words, it constructs the Amazon as a particular problem. Scholars studying the Brazilian Amazon often identify different governments throughout Brazilian history as proponents of a view of the region, its people, and forest [@drummond2006; @padua2012; @franchini2019; @capobianco2019; @pereira2021]. While the military dictatorship (1964-85) is, for example, associated with understanding the Amazon as issues of national sovereignty and economic integration, Sarney's (1985-1990) and Lula's (2003-2010) presidencies are often tied to environmental conservation. Yet, such monolithic representations advance the view that governments understood and acted consistently towards the Amazon as a specific problem, masking government diversity of discursive constructions and policies.

In this article, we show how the Brazilian Amazon has been constructed as a problem in presidential discourses since 1985. By adapting @bacchi2009's theory of problem-representation and @keck1998's theory of transnationalism (1998), we offer a framework that (1) explains transnational variation in problem-constructions as a function of speech setting and policies' outcomes, and (2) focuses on discourses rather than policies because they allow us to capture the diversity of ways policy goals are constructed to different audiences.

Many analyses of Brazilian environmental policies approach it as a national problem or an international problem, without a discussion of how important national actors behave differently in both scenarios. Our framework and methodology identify and accounts for this variation. We show, for instance, that when speaking to farmers in Cuiaba, one of Brazil's agricultural powerhouses, in a moment when deforestation is increasing, it is likely that any president would sidestep environmental issues, as President Lula did in the opening epigraph. Meanwhile, even President Bolsonaro at the United Nations General Assembly in New York City, spoke about protecting the Amazonian forests and its peoples.

Our focus on presidential speeches affords us analytical leverage and it is crucial to understand policy. The Brazilian executive branch of government, led by the president, has historically been the primary actor in shaping public policy in Brazil (Pinto 2016). While presidential speeches in Brazil have been studied for topics such as race relations [@silva2019] and inequality [@grangeia2017], the environment remains largely absent. Presidential speeches legitimize perceptions about the Amazon that are key for policy adoption and implementation [see @alesina2009; @lopez2020]. This is especially pertinent for deforestation in the Amazon as expectations of local actors, generated from material and discursive governmental practices, are crucial in decisions to deforest on the ground [@assuncao2015; @capobianco2019; @capobianco2021; @campbell2015].

Our findings are threefold. First, we find that the frequency in which the Amazon appears as a topic in presidential speeches is driven by domestic and international environmentally related events. Second, the same government adopts a diverse range of problem-construction. While the constructions of the Amazon as an issue of economic integration were more salient from 1985 to the mid-2000s, environmental conservation and social development problem-constructions steadily grew since the 2000s, temporarily surpassing economic integration as a more salient construction from 2010 to 2015. Constructing the Amazon as an issue of sovereignty steadily increased after 2010. Third, we find that presidents are more likely to construct the Amazon as an issue of environmental conservation when they speak at non-Amazonian countries outside of Brazil. Within the Amazon region and neighboring countries, presidents tend to construct the Amazon as an issue of social development and economic integration. This article provides the first comprehensive overview of when, how, and where the Amazon is constructed as problems in presidential speeches.

# 2 Theory: problem-constructions, transnational settings, and presidential speeches

## 2.1 Presidential speeches, policymaking, and problem-constructions

@@

There is more variation within a specific government, though, than suggested by the policy-cycle approach[@bacchi1999]. Political scientists have shown how governments have contradictory agrarian policies [@lopez2023], redistribution policies[@fairfield2017], health policies[@poncedeleon2021], and social policies[@niedzwiecki2017]. We argue that the reason behind these monolithic representations of governments in literature about the Amazon by social scientists is conceptual. As a key biome in global climate-mitigation efforts, the Amazon has often been conceptualized within the logic of a global collective action effort to halt environmental problems. Many studies of the Brazilian Amazon contain an explicit, or implicit assumption, that the problem is of environmental nature. If the Amazon is thought of as an "environmental" problem, as is often the case in the literature, then it demands an "environmental" solution. Consequently, more frequently than not, studies focus on understanding environmental policies [@capobianco2019], the work of environmental NGOs [@silva-muller2022], or the rise of transnational partnerships [@andonova2022]. We acknowledge the quality and importance of all these studies, but to build a comprehensive account of policies towards the region, we need a framework that assumes the possibility of varied problem-constructions beyond Amazonian environmental nature.

@@

@bacchi2009 [p. 2] places the question "what is the problem represented to be" in the center of policy studies . She argues that policies have a cultural dimension as they take "shape within specific historical and national or international contexts" @bacchi2009 [p. 10]. The existence or proposal of a policy implies that there is a (public) problem that needs (governmental) action to be fixed. The alleged problem is not always stated in a policy, but policies are represented as solutions for either explicit or implicit problems. Problem-representation matters for policy because it carries implications related to who is involved both in the root causes and in the solutions of the problem @bacchi2009. Following Bacchi, when speaking about a *policy object*, this is, a problem that demands dedicated policy attention, governments are either explicitly or implicitly constructing it as a problem of a certain kind. This is what we label as *problem-constructions*.\

Presidential discourse is a good instance to identify problem-constructions over time. While the relationship between discourse and policy may not be linearly causal, speeches have the power to introduce, define, and justify public policy, as well as shape issue perception to broad audiences [@zarefsky2004]. This is particularly true in the Brazilian context where the executive branch of government, led by the president, have historically been the primary players in shaping public policy, especially after the 1988 constitution [@pinto2016]. [more on presidents and their role in policies].

@@

The possibility of different problem-constructions for the same policy object in political discourse, requires an explanation of why they vary. Many analyses of Brazilian policies towards the Amazon approach it as a national problem or an international problem, without a discussion of how important national actors behave differently in both scenarios. By focusing on problem-constructions in presidential speeches, with an analytical strategy that distinguishes the location of the speech, we can move the empirical debate on the transnational nature of the Amazon forward. There are two main contenders for tying the problem-construction to international politics: Putnam's two-level game and Keck and Sikkink's transnational networks.

@putnam1988 theory of two-level games in international negotiations can be adapted to explain variation in presidential speeches. @putnam1988 argues that the outcomes of international negotiations lie within the overlap between the interests (and pressure) of domestic and international groups. At the national level, interest groups pressure the government to adopt favorable policies. Internationally, the government seeks to fulfill domestic preferences "while minimizing the adverse consequences of foreign developments" [@putnam1988, p. 434]. While the theory has been expanded to cover multi-levels games [see @collinson1999], its core remains the same: the outcome of negotiations should lie within the overlap of diverse interest groups.

@@

Putnam's theory of two-level games, though, is closer to a traditional view of policy, in which actors respond rationally to systemic pressures imposed by their position in relation to interest groups. Relatedly, it views summits as the UNFCCC as a purely intergovernmental body and thus of international nature. Its contribution, then, is to propose a connection between the national and international spheres as avenues for state-actors only. While there is some evidence of two-level games at play, this view seems incompatible with both the possibility of contradictory policies and of varied problem-constructions, which is more aligned with a constructivist perspective of politics. Rather than seeing international politics as the domain of only nation states, we start from the idea of transnationalism as relationships that transcends nation-states and incorporates non-state actors, spamming from the local level to the international level [@nye1971; @keck1998]. In this context, international summits are considered a transnational setting, as the presence and participation of non-state actors constitutes them.

@keck1998 theory of international politics as transnational networks, developed in the context of value-based advocacy groups, provides an alternative explanation for variation in problem-constructions. The authors reject the dichotomy of international and domestic levels arguing that the identity and goals of state and non-state actors are not derived from their structural location vis-a-vis domestic and foreign interest groups. Rather, international politics as transnational networks have "a structured and structuring dimension" [@keck1998, p. 4]: state and non-state actors participate in and shape international politics. In this sense, the positions state actors as components of a transnational network derive from complex interactions in which their identities and goals that are malleable and can vary across time and space.

While their theory is developed in the context of value-based advocacy groups, the overall view of international politics as structured by actors and structuring actors, should hold as well for state actors as well. Important participants of transnational networks, as presidents, have a myriad of tactics to introduce, define, persuade, socialize, and pressure when interacting with other actors. Transnational networks do not imply a consistent usage of problem-constructions across settings, as two-level games. That is, presidents construct problems and attempt to inform or leverage transnational audiences for symbolic or material gains according to their expectations of this audience. This indicates that how presidents construct policy objects as problems can vary, for instance, within the country ---speaking inside the Amazon versus outside--- because presidents identify these audiences as sources of different types of support. The same holds when speaking outside the country--- in Amazonian countries, non-Amazonian countries, and summits like Davos or UNFCCC--- where expectations and identities of audiences diverge completely from that of domestic actors.

This variation depends on the tactics presidents choose. Presidents can engage in information politics by creating politically meaningful information and/or moving it where they believe it can have an impact [@keck1998, p. 24]. In this regard, it is natural that presidents will always use policy adoption and outcomes in a cunning way, either to mask bad decisions or to boast successful ones. As well, presidents can engage in leverage politics by leveraging institutions to realize symbolic or material gains.  Material gains can be voters' support inside the Amazon or financial support in a UNFCCC Summit.

Our conceptual framework identifies problem-construction of the Brazilian Amazon in official presidential speeches and explains why and how it varies as a function of the setting where the speech takes place and policy outcomes. It advances the literature about the Amazon in two ways. First, we provide an outlook of the region, its people and forest beyond its environmental nature by focusing on different problem constructions as the main dependent variable of our study. In so doing, we avoid assuming what type of problem it is and focus on how it has been represented. Second, by analytically distinguishing transnational settings, one of our main independent variables, we avoid a single focus on the Amazon as either a national or an international problem.

@@

We start our research design by detailing how we operationalize the main variables of our theory. We then move to a section detailing our descriptive and inferential strategy. Along both sections, we explain the main limitations of our approach.

## 3.1 Data and operationalization strategy

To identify problem-constructions in presidential speeches, we need a dataset containing all official presidential speeches by Brazilian presidents. We build upon @cezar2020 dataset, which contains all speeches from 1985 to 2019. We update the dataset by scraping and adding all official speeches from 2020 to 2022. The final dataset encompasses 6240 speeches, with their dates and speech location. We proceed to identify speeches that refer to the Amazon as a region, people, or forest. We do so by detecting all speeches in which the stem "amazon" appears. In Portuguese, the stem captures terms such as "Amazonia", "Amazonica", "Amazonidas", "Amazonense(s)", "Amazonas", among others. We find that 968 speeches are, at least partially, about the Amazon. Using the poldis R package [@sposito2021], we extract two sentences before and after the sentence in which the stem "amazon" appears. We opt for picking two sentences around, rather than words, because sentences usually contain a cohesive idea. This is how our unit of analysis, an Amazonian statements, are generated. This process yields 2048 unique Amazonian statements. On average, an Amazonian statement contains 131 words.

### 3.1.1 Problem-construction in Amazonian Statements

To test our conceptual framework, we need to operationalize our main dependent variable, these are the possible problem-constructions for a policy object. To identify and classify Amazonian speeches as specific types of problem constructions, we developed a codebook inductively and with assistance of the literature. We identify three historical-political constructions: national sovereignty, economic integration, and environmental conservation. We argue a fourth separate one also exists: social development. The social development construction focuses on the constitutional rights and dignity of the inhabitants of the Amazonian region.

#### National Sovereignty

@@

#### Economic Integration

@@

#### Environmental Conservation

The rapid economic changes in the region from the 1960s to the 1980s were matched with the birth of environmental institutions [@drummond2006]. Arguably, the creation of these institutions stemmed from the impression of the lack of control over the environment following the efforts to economically integrate in the region [@acker2021; @capobianco2021; @hecht1990]. This process accelerated in the late 1980s with the birth of modern environmentalism [@viola1987], epitomized in the 1992 Earth Summit in Rio de Janeiro [@hochstetler2021; @capobianco2021]. The environmental conservation problem-construction emphasizes that the Amazon should be preserved, deforestation should be halted, and the sustainable practices of indigenous and local peoples should be maintained through the protection of their territories and rights to self-determination. The policy solution implies, for example, further investments in command-and-control institutions, the valuation of standing ecosystems through incentive schemes, and the creation of protected areas.

#### Social Development

@@

In their conceptualization, each problem-construction is mutually exclusive; that is, they cover different forms of constructing the Amazon as a problem. However, an Amazonian statement can be coded as referring to multiple problem constructions (see Figure 1). Amazonian statements, thus, can be pure-type statements (i.e. constructs the Amazon as a single problem) or mixed-types (i.e. constructs the Amazon as multiple issues).

@@

With the codebook in hand, each one of the authors separately hand-coded 1024 randomly selected Amazonian statements. This amount refers to 50% of all the Amazonian statements identified. We chose to hand code half of the observations because there are several nuances in how presidents talk about the Amazon over time. As well, since we have four different code categories, the size of the training set should be increased [see @grimmer2022]. The hand-coding strategy provides a robust and reliable number of observations to be used for training and validation of the automated model. Once hand coding was completed, the average intercoder agreement for hand-coded categories was 85%. For each non-matching coded observation, the authors discussed and sorted their disagreements, which were subsequently specified in the codebook.

Finally, the hand-coded data is randomly divided into a training set, containing 80% of the data (819 observations), and a validation set, containing the remaining 20% of the data (205 observations). We employ a support-vector machine (SVM) algorithm, a non-probabilistic linear classifier appropriate for binary categories, to label texts [@meyer2021; @noble2006]. To validate the model, the SVM algorithm is trained using the hand-coded training set and employed to classify observations in the validation set. As a final robustness check, the authors randomly select 100 automatically coded statements and find that the automated model is 95% accurate in coding these statements as the authors would. Automating the coding of the remaining half of the observations is estimated to have saved the authors over one month of work in comparison to manual coding. The final dataset for analysis, excluding false positive matches[^1], contains 1924 coded Amazonian statements.

[^1]: We also train the model to identify false positive matches in Amazonian statements. These are 124 Amazonian Statements that contain the stem "amazon", but are not necessarily meaningful as they might simply salute the governor of the Amazon, for example. We choose to remove these to reduce ambiguity and improve robustness of the findings.

One limitation of this approach is that our dataset covers only what is considered an official speech. Presidents, though, give interviews, appear in debates, talk at campaign rallies, and, more recently, post on social media. Problem-construction within presidential discourse, thus, happens in different public, private, and online settings for which we do not account or differentiate in this paper.

### 3.1.2 Operationalizing settings and policy outcomes

Our explanation for variation in problem-constructions is based on first the setting of a speech and second the policy outcomes.

#### Setting

Presidents construct problems and attempt to inform or leverage audiences for symbolic or material gains according to their expectations of the speech audience. To operationalize the audience, we extract the location of the speech and create four distinct settings. We argue that each one of these settings encompasses an audience that presidents can have different expectations and views, given the Brazilian Amazon is our policy object.

First, speeches can take place in Amazonian States inside Brazil. As a constituency with likely distinct demands than in other Brazilian states, presidents can try to communicate with them highlighting what presidents project their demands to be. Second, and consequently, Brazilian states outside the Amazon are another setting. In this case, as populations with some stake in the Amazon but not residing there, they might have different views of the Amazon.\

The third setting encompasses speeches that take place in Amazonian Countries, these are other countries that also harbor the Amazon Rainforest (e.g. Peru, Colombia, Venezuela, etc.). We argue that in bilateral visits to Amazonian countries, presidents can view other countries as potential peers in pushing specific visions of the rainforest internationally. The last setting encompasses all speeches in non-Amazonian countries. This entails speeches that take place within multilateral mechanisms as the UNFCCC, international summits as the World Economic Forum, bilateral visits to the United States or France, and so on. We argue that in these contexts, presidents also deploy problem-constructions aligned with what presidents perceive as possible expected gains from international processes.

While we would ideally control for the exact venue of the speech (multilateral mechanism, private fora, etc.), so we could have an even more precise identification of the audience, the location information of the speeches does not allow so. Nevertheless, we believe these four settings and their different audiences are already enough to demonstrate whether and how problem-construction indeed changes.

#### Policy outcome

We also expect presidents to have a cunning usage of policy outcomes. If outcomes are good, we argue that presidents will deploy aligned problem constructions to boast policy. If outcomes are bad, we argue presidents will mask them by highlighting a contradictory problem construction. We operationalize policy outcomes by using the yearly deforestation rates provided by INPE. We consider this a good option, as national and international actors do react to deforestation rates. In the late 1980s, for instance, there was strong media coverage in the USA of forest fires and deforestation figures. Similarly, in 2019, world press covered the August fires. We opt for lagging deforestation rates one year, as the consolidated numbers for 2007, for example, are published in 2008.

Many scholars inquire and find causal relationships between discourse and policy outcomes. While our research design and data does not allow us to test so, we do make arguments that might suggest a direction in certain parts....[further develop this limitation]

## 3.2 Data Analysis: descriptive and inferential methods

As a dataset like this was never constructed before, we start with longitudinal descriptive statistics, which can reveal interesting trends. First, we identify years in which Amazon as a general topic has been discussed more frequently than others and possible drivers. To do so, we take the share of yearly speeches that mention Amazon independently of the specific problem construction. Next, we discuss the evolution of yearly problem-constructions in time by showing our data longitudinally.We conclude with four separate fixed-effect logistic regression models that predict the probability of each problem-construction relative to the setting in which a speech takes place and the policy outcome.
